---
title: "Data Science and Analytics Midterm"
author: "Merve Çağlarer"
date: "08 12 2020"
output: html_document
---

# Online Shoppers' Intention

## Introduction

### We are using the “Online Shoppers Intention” (https://archive.ics.uci.edu/ml/datasets/Online+Shoppers+Purchasing+Intention+Datase) dataset related to visits of customers to an online store and their decision on purchase a products.The goal of the project is to analyze by using unsupervised learning algorithms and to make predictive models around online shoppers’ intentions.

### The dataset consists of 18 features which are 10 numerical and 8 categorical belonging to 12,330 sessions.

#### -Administrative: Number of pages visited by the visitors about account management.
#### -Administrative Duration: Total amount of time (in seconds) spent by the visitors on account management related pages.
#### -Informational: Number of pages visited by the visitors about Web site, communication and address indormation of the shopping site.
#### -Informational Duration: Total amount of time (in seconds) spent by the visitors on informational pages.
#### -Product Related: Number of pages visited by the visitors about product related pages.
#### -Product Related Duration: Total amount of time (in seconds) spent by the visitors on product related pages.
#### -Bounce Rate: This is the number of single-page visits by visitors of the website.
#### -Exit Rate: This is the number of exits from the website.
#### -Page Value: It represents the average value for a web page that a user visited before completing an e-commerce transaction. It tells you which specific pages of the site offer the most value.
#### -Special Day: The closeness of the site visiting time to a specific special day (e.g. Mother’s Day, Valentine’s Day).
#### -Operating Systems: Operating system of the visitors.
#### -Browser: Browsers of the visitors.
#### -Region: Geographic region from which the sesssion has been started by the visitors.
#### -Traffic Type: Traffic source by which visitor has arrived at the Web site (e.g. banner, SMS, direct).
#### -Visitor Type: Visitor type as "New Visitor", "Returning Visitor", "Other".
#### -Weekend: Boolean value indicating whether the date of the visit is weekend or not.
#### -Month: Month value of the visit date.
#### -Revenue: Class label indicating whether the visit has been finalized with transaction.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import the relevant libraries.

```{r libraries}

library(plyr)
library(dplyr)
library(ggplot2)
library(scales)
library(lubridate)
library(tidyverse)
library(gridExtra)
library(openxlsx)
library(factoextra)

```

## Load the data and convert characteristic data to numerical univariate. 

```{r load_data}

online_shoppers <- read.csv("C:/Users/Merve/Downloads/DataScience/r codes/online_shoppers_intention.csv")
summary(online_shoppers)

```

#### There are duplicate rows in the dataset, after removing 125 duplicates, 12205 records remain and there is no missing value.

```{r clean_data}

str(online_shoppers)
duplicate_number <- nrow(online_shoppers[duplicated(online_shoppers),])
online_shoppers <- online_shoppers[!duplicated(online_shoppers),]
str(online_shoppers)
sum(is.na(online_shoppers))
summary(online_shoppers)

```

## Exploratory Data Analysis

```{r convert_data}

online_shoppers <- transform(online_shoppers,
                  Month = as.factor(mapvalues(Month, 
                                               c("Feb","Mar","May","June","Jul","Aug","Sep","Oct","Nov","Dec"),
                                               c(2,3,5,6,7,8,9,10,11,12))),
                  VisitorType = as.factor(mapvalues(VisitorType,
                                                    c("Returning_Visitor","New_Visitor","Other"),
                                                    c(1,2,3))),
                  Weekend = as.factor(mapvalues(Weekend,
                                                c("TRUE","FALSE"),
                                                c(1,0))))

```

### Let's show why the feature is unnecessary for analysis.

#### Administrative

```{r Administrative}

g <- ggplot(online_shoppers, aes(Administrative))
g + geom_density(aes(fill=factor(Weekend)), alpha=0.8) + 
  labs(title="Density plot", 
       subtitle="Number of account management pages visited by the Weekend",
       x="Administrative",
       fill="Weekend")

```

##### This graph shows that the user usually visits the page at the same rate on weekends or weekdays. So 'Administrative' is a convenient feature to extract from data. 

#### Administrative Duration

```{r Administrative_Duration}

g <- ggplot(online_shoppers, aes(Administrative_Duration, Weekend))
g + geom_jitter(width = .5, size=1, colour="blue") +
  labs(subtitle="Total amount of time spent in account management pages visited by the Weekend", 
       y="Administrative_Duration", 
       x="Weekend")

```

##### This graph shows that the user usually visits the page at the same duration rate on weekends or weekdays. So 'Administrative_Duration' is a convenient feature to extract from data. 

#### Informational

```{r Informational}

g <- ggplot(online_shoppers, aes(Informational))
g + geom_density(aes(fill=factor(Weekend)), alpha=0.8) + 
  labs(title="Density plot", 
       subtitle="Number of information pages visited by the Weekend",
       x="Informational",
       fill="Weekend")

```

##### This graph shows that the user usually visits the page at the same rate on weekends or weekdays. So 'Informational' is a convenient feature to extract from data. 

#### Informational Duration

```{r Informational_Duration}

g <- ggplot(online_shoppers, aes(Informational_Duration, Weekend))
g + geom_jitter(width = .5, size=1, colour="red") +
  labs(subtitle="Total amount of time spent in information pages visited by the Weekend", 
       y="Informational_Duration", 
       x="Weekend")

```

##### This graph shows that the user usually visits the page at the same duration rate on weekends or weekdays. So 'Informational_Duration' is a convenient feature to extract from data. 

#### Product Related

```{r Product_Related}

online_shoppers %>% 
  ggplot() +
  aes(x = ProductRelated) +
  geom_bar() +
  facet_grid(Weekend ~ .,
             scales = "free_y")

```

##### This graph shows that the user usually visits the page at the same rate on weekends or weekdays. So 'Product_Related' is a convenient feature to extract from data. 

#### Product Related Duration

```{r ProductRelated_Duration}

g <- ggplot(online_shoppers, aes(ProductRelated_Duration, Weekend))
g + geom_jitter(width = .5, size=1, colour="green") +
  labs(subtitle="Total amount of time spent in product pages visited by the Weekend", 
       y="ProductRelated_Duration", 
       x="Weekend")

```

##### This graph shows that the user usually visits the page at the same duration rate on weekends or weekdays. So 'ProductRelated_Duration' is a convenient feature to extract from data. 

#### Bounce Rate

```{r BounceRate}

plot(online_shoppers$BounceRates,online_shoppers$Weekend,xlab = "BounceRates",ylab="Weekend")

```

#### Exit Rate

```{r Exit_Rate}

plot(online_shoppers$ExitRates,online_shoppers$Weekend,xlab = "ExitRates",ylab = "Weekend")

```

#### Page Value

```{r PageValue}

plot(online_shoppers$PageValues,online_shoppers$Weekend,xlab = "PageValues",ylab = "Weekend")

```

##### These 3 graphs are also shown that, there is not enough affect for exploraty. We should do PCA for them.

##  Segmentation

### I want to analyze visitor status on weekend == 1.

```{r Month_Weekend}

Month_Weekend<- online_shoppers %>% filter(online_shoppers$Weekend==1)
ggplot(Month_Weekend, aes(Month_Weekend$Month, ..count..)) + 
    geom_bar(aes(fill = Month_Weekend$Weekend), position = "dodge") + 
  labs(x="Month",y="Weekend")

```

##### This graph shows that the user visits the shopping page the most 11. month which is November and after that 5. month which is May in on weekends.

```{r VisitorTypes_Weekend}

VisitorTypes_Weekend <- online_shoppers %>%
  filter(Weekend == 1 ) %>%
  select((VisitorType:Weekend))


ggplot(data.frame(VisitorTypes_Weekend), aes(y = VisitorType, col = Weekend, fill = Weekend)) + geom_bar() + 
 labs(title = "VisitorType - Weekend Bar", y = "VisitorType")

```

#####  This graph shows that the Returning_Visitor visits the shopping page on weekends, after that New_Visitor and Other.

## PCA Analysis

#### First of all, we have to convert our data to numerical.

```{r PCA_all}

sapply(online_shoppers, class)
online_shoppers$Administrative = as.numeric(as.factor(online_shoppers$Administrative))
online_shoppers$Informational = as.numeric(as.factor(online_shoppers$Informational))
online_shoppers$ProductRelated = as.numeric(as.factor(online_shoppers$ProductRelated))
online_shoppers$Month = as.numeric(as.factor(online_shoppers$Month))
online_shoppers$OperatingSystems = as.numeric(as.factor(online_shoppers$OperatingSystems))
online_shoppers$Browser = as.numeric(as.factor(online_shoppers$Browser))
online_shoppers$Region = as.numeric(as.factor(online_shoppers$Region))
online_shoppers$TrafficType = as.numeric(as.factor(online_shoppers$TrafficType))
online_shoppers$VisitorType = as.numeric(as.factor(online_shoppers$VisitorType))
online_shoppers$Weekend = as.numeric(as.factor(online_shoppers$Weekend))
online_shoppers$Revenue = as.numeric(as.factor(online_shoppers$Revenue))

#PCA for all clusters
data_all <- online_shoppers[,apply(online_shoppers, 2, var, na.rm=TRUE) != 0 ]
pca_all<- prcomp(data_all, center = TRUE, scale. = TRUE)
summary(pca_all)

```

##### We have 18 components for 18 features. First 3 variances are 0.193 0.1036 0.08084 and First 3 Cumulative Proportions are 0.193 0.2965 0.37735.This means that if we use 3 PCAs, we will get 37% information with the 3 dimentionals dataset.We should take components that we can preserve at least 70-80 % of the cumulative variance.

```{r PCA}

#I put the data into what I wanted to see how the weekend affected.
shoppers <- online_shoppers %>%
  filter(Weekend == 1)%>%
  select(SpecialDay:Revenue)

sapply(shoppers, class)

#PCA for 8 clusters
data <- shoppers[,apply(shoppers, 2, var, na.rm=TRUE) != 0 ]
pca <- prcomp(data, center = TRUE, scale. = TRUE)
pca
summary(pca)

```


##### We have 8 components for 8 features.The first 3 Variances are 0.1935 0.1626 0.1243 and the first 3 Cumulative Proportions are 0.1935 0.3561 0.4804.This means that if we use 3 PCAs, we will get 48% information with the 3 dimentionals dataset.


```{r PCA_plot1}

plot(pca, main = "Principal Components")
screeplot(pca, type = "l", main = "Screeplot of the first 4 PCs")
abline(h = 1, col="red", lty=5)
legend("topright", legend=c("Eigenvalue = 1"),
       col=c("red"), lty=5, cex=0.6)

```

##### In additon, Eigen value shows us, 3 dimentions are enough for our dataset from the plot.

```{r PCA_plot2}

fviz_pca_var(pca,
             col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE)

```

##### This graph shows that, Browser, Region, TrafficType, VisitorType, OperatingSystems, Weekend, Month, SpecialDay affect the data in the same direction.

```{r PCA_plot3}

revenue <- as.factor(shoppers$Revenue)
fviz_pca_ind(pca, geom.ind = "point", pointshape = 21, 
             pointsize = 2, 
             fill.ind = revenue, 
             col.ind = "black", 
             palette = "jco", 
             addEllipses = TRUE,
             label = "var",
             col.var = "black",
             repel = TRUE,
             legend.title = "Revenue") +
  ggtitle("2D PCA-plot with 9 features") +
  theme(plot.title = element_text(hjust = 0.5))

```

##### This graph shows that, when cluster processing with Revenue variables are done, at least close results could be obtained.

## Clustering

### K-means 

```{r Kmeans2}

#Kmeans with 2 clusters
km2 <- kmeans(shoppers[,-ncol(shoppers)], 2)

table(km2$cluster, shoppers$Revenue)

```

##### Revenue==True and Revenue==False are mostly in 2. cluster.

```{r Kmeans_plot}

plot(pca$x[,1],pca$x[,2],
     xlab="PC1", ylab = "PC2", 
     main = "PC1 / PC2 - plot", 
     col = km2$cluster)

```

##### Base on the similarities, we have 2 clusters. 

```{r Kmeans3}

#Kmeans with 3 clusters
km3 <- kmeans(shoppers[,-ncol(shoppers)], 3)

table(km3$cluster, shoppers$Revenue)

```

##### Revenue==True is mostly in 1. cluster and Revenue==False is mostly in 3. cluster.This shows that 3 cluster is not suitable for our dataset.

##  Hierarchical Clustering

```{r  hierarchical}

distance <- dist(shoppers[1:100,], method="euclidean") 
hier <- hclust(distance, method="average")
plot(hier, cex=0.7) 
rect.hclust(hier, k=2, border="red")

```

##### The dendrogram given below is the output of hierarchical clustering for the first 100 visitors. Average link method and Euclidian distance was used to generate the plot. 

```{r hierarchical2}

distance <- dist(shoppers[1:100,], method="euclidean") 
hier <- hclust(distance, method="average")
plot(hier, cex=0.7) 
rect.hclust(hier, k=3, border="red")

```

##### The dendrograms also show that 2 clusters is more suitable than 3 clusters.

## Conclusion
##### In this project, we explore how costomers' purchasing decisions change based on 18 characteristics. For example, we generally believe that "SpecialDay" has a strong impact on the decision, as we have more free time to browse online shopping on weekends. In addition, visitor types could vary depending on whether it is a weekend or not. As a result, when I examine the online shopping movements of the visitors from the data using PCA, Kmeans and Hierarchical Clustering; the result of our model shows that the weekend will not affect much.
